Fish do not need Visas: the transboundary nature of the world’s exploited marine fish species

### 2019-12-24 Manuscript_Draft.Rmd
Included RFMO analysis

Distributions of marine species around the world are not bounded by human-made boundaries; rather shaped by contemporary biotic and abiotic factors, as well as their evolutionary history [@Hunch; @NelsonFishesOfTheWorld]. However, management of living marine resources is predicated on the definition of “stocks” that are delineated by human-made spatial boundaries that often do not correspond to biologically-meaningful population units [@Song:2017iua]. The delineation of Economic Exclusive Zones (EEZs) under the United Nations Convention on the Law of the Seas (UNCLOS) in the early 80s, for instance, established political boundaries across the distribution of many species, creating shared species between nations. Theories and empirical experience demonstrate fisheries targeting resources that straddle between political boundaries complicate fisheries management and potentially reduce its effectiveness to achieve stated objectives.. For example, fishing quotas set by different nations for a shared fishery resource have led to conflict [@Spijkers:2017ij]. Climate change, which is shifting species distributions, is likely to exacerbate this type of conflict [@Pinsky:2018cb] and presents further challenges for fisheries management [@Mills].

Forty years after the formal adoption of UNCLOS, accurate estimates of the number of exploited marine species that distribute across two or more EEZs (hereafter called shared species) in the world is still unclear. Caddy in 1997 estimated that there could be up to 1,500 transboundary fish species in the world. However, such estimation lacked a proper assessment due to limited information on the distribution of species at the time [@Caddy:1997ue]. More recently, Teh et al. (2010) estimated that 206 species can be considered shared, accounting for a total catch of 34.2 million t and a global landed value of USD 30.7 trillion. While this represents a good starting point, the analysis was based on a literature review of shared species and did not include a mechanistic way of determining the transboundary nature of the species [@Teh:2015gd]. Getting accurate estimates of the number of shared species and where they are located can help develop fisheries management and conservation plans that account for the challenges from transboundary resources management [@Millis].

@@

This paper aims to identify the world current transboundary species and determine their contribution to the catch and revenue of the world marine fisheries. The Food and Agricultural Organization (FAO) recognizes three types of shared stocks: (i) transboundary stocks, shared by 2 or more neighboring coastal nations; (ii) straddling stocks, occurring in two or more adjacent national jurisdictions and the high seas; and (iii) highly migratory stocks, found in the EEZs of coastal nations that are not necessarily adjoining, and the high seas. We focus on estimating the number of species that meet the first category. We overlay the known distribution of **968** commercially valuable marine fish species and the Exclusive Economic Zones of **175** coastal countries to determine the total number of transboundary species and their contribution to global catch and landed value. We carry out the analysis at the species level, rather than stocks (populations within a species), due to lack of stock-specific spatial and ecological information on all of the fished species [@Teh:2015gd].

We used four-criteria that combined multiple species distribution data-sources and the world's EEZs to determine if a species can be considered a “shared species”. The first criteria only considered shared species between neighbouring EEZs, rather than the extended distribution. We sub-divided EEZs by regions (e.g. Mexico Pacific and Mexico Atlantic). Current species distributions were predicted from occurrence records obtained from publicly available observational databases of the world’s marine species [@REFNEreus] and two structurally distinct species distribution models [@REFGabs; @Zeller] (see Methods). For the second criteria we considered a species to be present in a grid cell if all data sources predicted positive occurrence. The third criteria determined whether the species were exploited at the grid cell using spatially explicit catch data [@Zeller]. A species was kept if both catch data and predicted species occurrences were positive at the same spatial cells. Finally, the last criteria determined a spatial treshold to confirm that a species occurs in both neighboring EEZs. Once we have identified the shared stocks, we estimated the contribution of the species to fishing revenues and landings using the SAU database [@Zeller] and categorized them according to catch trends [@Kleisner:2013gh].

@@

Overall, we estimated **633** transboundary species in the world, more than double the number previously estimated [@Teh:2015gd], accounting for **65.4**% of the species analyzed (Fig. 1A). Between 2005 and 2014, national fleets targeting shared stocks within EEZs landed an average of **69.2** million tonnes, representing **64**% of the EEZ-global catch, and generating USD **79.7** million in fishing revenue. This is equivalent to **49.8**% of the global revenue of the same period. These values are higher than previously estimated reinforcing the importance of these stocks at the global level [@Teh].

Transboundary species are particulary important for Northern America (mean = 3,901 $\pm$ 4,497 million USD) and Eastern Asia (mean = 2,827 $\pm$ = 2,667 million USD), significantly differing from the rest in tn terms of revenue (one-way ANOVA; *F*(16,157) = 4.91, *p* < 0.001, $\alpha$ = 0.05) (Fig 1B). The large variation in revenue from these regions is driven by China (USD 6,279 million), in E. Asia and the US (USD 9,605 million) in N. America who contribue 55% and 83% of the region's revenue, respectiveley. Russia (USD 6,361 million), Peru (USD 5,211) and Japan (USD 3,369) complete the top five countries with more revenue from transboundary species.

Catches, however, are only significantly different in E. Asia (one-way ANOVA; *F*(16,155) = 3.341, *p* < 0.001, $\alpha$ = 0.05), with China, the world's top fish producer with 15.3 million tons landed in 2016 [@SOFIA2018], getting around 1/3 of its capture from transboundary species. For the US, 16 species make up 90% of the revenue from transboudary species, with American scallop (*Placopecten magellanicus*) shared with Canada in Georges Bank [@https://elischolar.library.yale.edu/cgi/viewcontent.cgi?article=1008&context=fes-pubs] contributing 41.8% or the revenue. Russia shares the largest fished species in the world, Alaskan polllock (*Theragra chalcogramma*) [@SOFIA2018], with both Japan and the US. Rhis fishery is responsible for 50% of the Russian revenue from transboundary species, with other eight species comprising the 90% of revenue. In the case of Peru, Peruvian anchoveta (*Engraulis ringens*) and jumbo squid (*Dosidicus gigas*), both shared with Chile, represent almost 90% of the revenue from transboundary species, with anchoveta accounting for 79%.

@@

We found transboundary fisheries to be of substantial economic importance in regions with past fisheries-related conflicts [@Spijkers:2019cz]. For example, the United States and Canada had a dispute over Pacific salmons (*Oncorhynchus sp*) that lasted over 10 years and was only resolved after these two nations set a series of regulations to compensate losses [@Miller2003]. Since 2007, the European Union, Norway, Iceland, and the Faroe Islands have been at odds over the size and relative allocation of total allowable catches (TACs) for mackerel [@Spijkers:2017ij], a species that is transboundary and contributes **X** to landings and **X** in revenue to the countries involved in the dispute. In addition, climate change could exacerbate conflicts over transboundary species as climate change shifts the distribution of marine species around the globe [@Pinsky].

There is no significant difference in the exploitaion rate of transboundary species between continents (one-way ANOVA *F*(5,443), *f* = 1.24, *p* = 0.289, $\alpha$ = 0.05) as most species are fished over 50% of their historical maximum peak (Category C) in all continents, but Oceania (Fig.2). However, the number of species exploited by each country in each category significantly differs from each other (one-way ANOVA *F*(5,442), *f* = 322.9, *p* < 0.001, $\alpha$ = 0.05). Overall, most transboundary species are fished over 50% of their once maximum peak (Category C) suggesting that their current exploitation rate is over the maximum sustainable yield [@Kleisner:2011wn].

Empirical analysis suggest that cooperation will have better outcomes for shared fisheries management. Examples vary between regions and species including Mexico and the United States over Pacific anchovies [@CisnerosMontemayor:2020kv], Norway and Russia over Atlantic cod (*Gadus morhua*)[@Diekert:2010gp], and between Namibia and South Africa over hake (*Merluccius spp*) [@Sumaila:2003vw].

Indeed, there are some successful bilateral management plans for species such as Pacific Halibut between Canada and the US as well [@REference?] and multilateral treaties like the Pacific Nauru Agreement that manages 30% of global skipjack tuna in the South Pacific Islands [@Aqorau:2018bh]. However, the effectiveness of Regional Fisheries Management Organizations, responsible for managing high migratory stocks in areas beyond national jurisdiction, has been previously questioned [@CullisSuzuki:2010fi]. Countries that are today harvesting species at rates less than 50% of the historical maximum catch (Category C) should be paying extra care in these species (Fig. 2). While these catch trends do not reflect stock status [@Branch:2011br], they could reassemble fisheries that are over exploited or collapsed [@Kleisner:2011wn]. The rebuilding of a transboundary species could be strongly impacted by uncoordinated efforts of countries harvesting the same species but with different objectives [@Ref].

@@

Species can have a cosmopolitan distribution, that is, a broader distribution, or be limited to a specific region, known as endemic [@REF]. Not surprisingly, highly migratory pelagic-oceanic species such as skipjack tuna (*Katsuwonus pelamis*), Common thresher shark (*Alopias vulpinus*), and tunas (*Thunnus sp.*) are among the more shared resources in the world with a median of 40 nations sharing these resources (Fig 3.). These species are very important in terms of economic value as they (global) average revenue exceeds USDXXX.

The median for species of all other ecosystem preferences is close to, or less than, 20, as many of these might be cosmopolitan but not necessarily highly migratory (e.g. *Solea solea*, *Panulirus argus*) (Fig. 3). However, less distributed species must not be ignored. The world of fisheries is highly connected trough larvae dispersal [@Ramesh:2019va] and so, even if a population swims between the borders of two nations, their larvae could be supplying populations thousands of kilometers away [@Ramesh:2019va].

Moreover, many coastal communities, specially in developing regions, and highly dependent on fishing local fish (e.g. reef fish) for food security and improved livelihoods [@CisnerosMontemayor:2016gq; Wabnitz:2018gf].

@@

Management of transboundary fisheries needs to be looking forward. As species shift their known distributions to cope with a changing ocean, new transboundary species will arise, challenging the management of shared stocks worldwide [@Pinsky]. Conflicts in regions with a high abundance in transboundary species will likely be exacerbated by climate change, as seen in the ongoing dispute between the European Union, Norway, Iceland and the Faroe Islands over mackerel (**species**) [@Spijkers:2017ij]. Understanding how climate change will affect the biogeography of marine species, and notably the number of countries' EEZ it spans is a key element to the design of effective long-term spatial-based fisheries management [@Alexa, @Song:2017iua].

@@

<!-- Overview of West African fisheries under climate change: Impacts, vulnerabilities and adaptive responses of the artisanal and industrial sectors -->
<!-- https://www.sciencedirect.com/science/article/pii/S0308597X16302901#bib4 -->

@@

The ENM-Nereus consists in a multimodel approach based on a Bioclim and a Bososted Regression Tree model [@Thuiller:2009gp], a Maxent model [@Phillips:2006ff], and a Non-Parametric Probabilistic Ecological Niche Model [@Beaugrand:2011fd]. Environmental variables utilized in the models included sea surface temperature, surface pH, surface oxygen concentration, and vertically integrated (0–100 m) net primary production (NPP) [@Asch:2018ca]. The ENM-Nereus employed all of the observational data-sources previously mentioned. Global environmental conditions were obtained and results averaged from three Earth System Models developed by the Geophysical Fluid Dynamics Laboratory (GFDL- https://www.gfdl.noaa.gov/earth-system-model/), the Institute Pierre Simon Laplace (IPSL- www.icmc.ipsl.fr/), and the Max Planck Institute for Meteorology (MPI- www.mpimet.mpg.de/en/science/models/). See [@Asch:2018ca; @Reygondeau:2019uh] for model details.

@@

The previous models combine observational data with a series of biotic and abiotic information to determine the probability that a species will be found in a given space at a given time. However, this does not mean that the species in question will actually be there. While the models do use approaches to double-check species occurrences (e.g. ENM-Nereus uses four different species distribution algorithms and SPD-SAU undertakes validation by means of other models), we used a fourth data source to corroborate the models’ outputs. The *Sea Around Us* estimates total reconstructed catches - i.e., catches  based on all publicly available information sources and including discards, as well as unreported and illegal catches that are not included in the FAO data available for each country. Catches are also spatially allocated [@Zeller2016]. Thus, we used the Sea Around Us catch reconstruction database from 2005 to 2014 (last ten yeares of data) as the fourth dataset to estimate transboundary species and to estimate their catch contribution within EEZs (average 2005-2014).

@@

We developed a four-criteria method for determining whether or not a species was transboundary. Only species that meet all criteria were considered as transboundary species and the analysis was done only within the EEZs of coastal states. In some cases, criteria also work as indexes to measure the uncertainty in the analysis as they provided continuos index (e.g. from 0 to 1) rather than a discrete, yes or no, index . All of the analysis was done in the statistical software *R version 3.5.2 (2018-12-20)* with the pacages *tidiverse*[@],*data.table*[@],*wesanderson*[@],*janitor*[@],*spdep*[@], *sf*[@],*sp*[@],*rgdal*[@],*tools*[@],*parallel*[@],*rfishbase*[@],*zoo*[@]. All code is available at https://github.com/jepa/FishForVisa.

@@

We assume that a species was only present in a given grid-cell if it was reported in the SAU catch database. Therefore, all species that were not reported as caught in any single year between the reference years (1970 to 2000) in a given grid-cell were dropped, regardless of the *Species Index*. The assumption relies in that if a commercial species is projected within any fishing country, such species would have been fished, and thus reported at some period of time, thus validating the models.

@@

We estimated the economic contribution of transboundary species for each country using global ex-vessel price [@Tai2017]. The data includes ex-vessel price from multiple sources and a structured interpolation method (e.g. similar countries, species) to fill in data gaps [@Sumaila:2015uc]. The contribution of transboundary species was done by aggregating the catch of each species that happened within the EEZ of the countries sharing that species. For this study, we did not include catch from areas beyond national jurisdiction. Finally, we use the catch data to determine the exploitation category of each species within a EEZ. Although this method has previously used to estimate stock status [@Grainger:1996tk; @Kleisner:2011wn], the categories presented here are intended to be seen as catch trends, and not status of each species. We only assessed species within each EEZ that had at least ten years of data between the first and last reported landings with at least five consecutive data years. Three categories were drawn from the method depending on the year's catch, the maximum catch over time and the minimum catch after the catch peak (**Table XX**). Finally, we reported the predominant category over the last ten years of catch data (2004-2014).

@@

# Recicled text (Not to be included)

>>>>

Distributions of marine species around the world are not bounded by human-made boundaries; rather shaped by contemporary biotic and abiotic factors, as well as their evolutionary history [@Hunch; @NelsonFishesOfTheWorld]. However, management of living marine resources is predicated on the definition of “stocks” that are delineated by human-made spatial boundaries that often do not correspond to biologically-meaningful population units [@Song:2017iua]. The delineation of Economic Exclusive Zones (EEZs) under the United Nations Convention on the Law of the Seas (UNCLOS) in the early 80s, for instance, established political boundaries across the distribution of many species, creating shared species between nations. Theories and empirical experience demonstrate fisheries targeting resources that straddle between political boundaries complicate fisheries management and potentially reduce its effectiveness to achieve stated objectives [@REF]. For example, evironmental fluctuation leading to changes in the stock distribution [@Mills] and fishing quotas set by different nations for a shared fishery resource have led to conflict [@Spijkers:2017ij]. Moreover, climate change, which is shifting species distributions, is likely to exacerbate this type of conflict [@Pinsky:2018cb] and presents further challenges for fisheries management [@Mills].

Forty years after the formal adoption of UNCLOS, accurate estimates of the number of exploited marine species that distribute across two or more EEZs (hereafter called shared species) in the world is still unclear. Caddy in 1997 estimated that there could be up to 1,500 transboundary fish species in the world. However, such estimation lacked a proper assessment due to limited information on the distribution of species at the time [@Caddy:1997ue]. More recently, Teh et al. (2010) estimated that 206 species can be considered shared, accounting for a total catch of 34.2 million tones and a global landed value of USD 30.7 trillion. While this represents a good starting point, the analysis was based on a literature review of shared species and did not include a mechanistic way of determining the transboundary nature of the species [@Teh:2015gd]. Getting accurate estimates of the number of shared species and where they are located can help develop fisheries management and conservation plans that account for the challenges from transboundary resources management [@Millis].

@@

This paper aims to identify the world current transboundary species and determine their contribution to the catch and revenue of the world marine fisheries. The Food and Agricultural Organization (FAO) recognizes three types of shared stocks: (i) transboundary stocks, shared by 2 or more neighboring coastal nations; (ii) straddling stocks, occurring in two or more adjacent national jurisdictions and the high seas; and (iii) highly migratory stocks, found in the EEZs of coastal nations that are not necessarily adjoining, and the high seas. We focus on estimating the number of species that meet the first category. We overlay the known distribution of **968** commercially valuable marine fish species and the Exclusive Economic Zones of 153 coastal countries to determine the total number of transboundary species and their contribution to global catch and landed value. We carry out the analysis at the species level, rather than stocks (populations within a species), due to lack of stock-specific spatial and ecological information on all of the fished species [@Teh:2015gd].

We used four-criteria that combined multiple species distribution data-sources and the world's EEZs to determine if a species can be considered a “shared species”. The first criteria consisted in only considering shared species between neighbouring EEZs, rather than the extended distribution. For the second criteria we considered a species to be present in a grid cell if all data sources predicted positive occurrence. Current species distributions were predicted from occurrence records obtained from publicly available observational databases of the world’s marine species [@REFNEreus] and two structurally distinct species distribution models [@REFGabs; @Zeller] (see Methods). The third criteria determined whether the species were exploited at the grid cell using spatially explicit catch data [@Zeller]. A species was kept if both catch data and predicted species occurrences were positive at the same spatial cells. Finally, the last criteria determined a spatial treshold to confirm that a species occurs in both neighboring EEZs. Once we have identified the shared stocks, we estimated the contribution of the species to fishing revenues and landings using the SAU database [@Zeller] and categorized them according to catch trends [@Kleisner:2013gh].

@@

Overall, we estimated **633** transboundary species in the world, more than double the number previously estimated [@Teh:2015gd], accounting for **65.4**% of the species analyzed (Fig. 1A). Between 2005 and 2014, national fleets targeting shared stocks within EEZs landed an average of **69.2** million tonnes, representing **64**% of the EEZ-global catch, and generating USD **79,745** million in fishing revenue. This is equivalent to **49.8**% of the global revenue of the same period. These values are higher than previously estimated reinforcing the importance of these stocks at the global level [@Teh].

@@

Transboundary species are particulary important for Northern America (mean = 3,901 $\pm$ 4,497 million USD) and Eastern Asia (mean = 2,827 $\pm$ = 2,667 million USD), significantly differing from other regions in terms of revenue (one-way ANOVA; *F*(16,157) = 4.91, *p* < 0.001, $\alpha$ = 0.05) (Fig 1B). The large variation in revenue from these regions is driven by China (USD 6,279 million), in E. Asia and the US (USD 9,605 million) in N. America who contribue 55% and 83% of the region's revenue, respectiveley. Russia (USD 6,361 million), Peru (USD 5,211) and Japan (USD 3,369) complete the top five countries with more revenue from transboundary species. These 5 nations are responsable for 30% of the global revenue from transboundary species.

Catches, however, are only significantly different in E. Asia (one-way ANOVA; *F*(16,155) = 3.341, *p* < 0.001, $\alpha$ = 0.05), with China, the world's top fish producer with 15.3 million tons landed in 2016 [@SOFIA2018], getting around 1/3 of its capture from transboundary species. For the US, 16 species make up 90% of the revenue from transboudary species, with American scallop (*Placopecten magellanicus*) shared with Canada in Georges Bank [@https://elischolar.library.yale.edu/cgi/viewcontent.cgi?article=1008&context=fes-pubs] contributing 41.8% or the revenue. Russia shares the largest fished species in the world, Alaskan polllock (*Theragra chalcogramma*) [@SOFIA2018], with both Japan and the US. This fishery is responsible for 50% of the Russian revenue from transboundary species, with other eight species comprising 90% of the revenue. In the case of Peru, Peruvian anchoveta (*Engraulis ringens*) and jumbo squid (*Dosidicus gigas*), both shared with Chile, represent almost 90% of the revenue from transboundary species, with anchoveta accounting for 79%.

The exploitation category of transboundary species significantly differs from each other (one-way ANOVA *F*(5,446), *f* = 324.5, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05 for all) with most species fished under 50% of their historical maximum peak (Fig.2). This is true for all (but Oceania) continents with no significan difference between them (one-way ANOVA *F*(5,443), *f* = 1.24, *p* = 0.289, $\alpha$ = 0.05), suggesting that the current global exploitation rate of transboundary species could be over the maximum sustainable yield [@Kleisner:2011wn; @PaulyNature2013]. Countries that are today harvesting at this rate (Category C) should be paying extra care in the management of these species (Fig. 2).

@@

While these catch trends do not directly reflect stock status [@Branch:2011br] - e.g. other factors like markets, fishing policies or extreem events also influence catch volume [@PaulyNature2013] - they could ressemble fisheries that are over exploited or collapsed [@Kleisner:2011wn]. Empirical analysis suggest that management of transboundary species will have better outcomes in terms of fish yield and sustainabillity whith cooperation between nations [@REF]. Examples vary between regions and species including Mexico and the United States over Pacific anchovies [@CisnerosMontemayor:2020kv], Norway and Russia over Atlantic cod (*Gadus morhua*) [@Diekert:2010gp], and Namibia and South Africa over hake (*Merluccius spp*) [@Sumaila:2003vw].

Collaboration can also reduce conflict between nations over shared ressources. Transboundary fisheries are abundant and of economic importance for some of the top countries with past fisheries-related conflicts like Canada, the US, China and Russia [@Spijkers:2019cz]. For example, the United States and Canada had a dispute over Pacific salmons (*Oncorhynchus sp*) that lasted over 10 years and was only resolved after these two nations set a series of regulations to compensate losses [@Miller2003]. Since 2007, the European Union, Norway, Iceland, and the Faroe Islands have been at odds over the size and relative allocation of total allowable catches (TACs) for mackerel [@Spijkers:2017ij], a species that is transboundary and contributes **X** to landings and **X** in revenue to the countries involved in the dispute. In addition to past and present conflicts, climate change could exacerbate conflicts over transboundary species as climate change shifts the distribution of marine species around the globe [@Pinsky].

First steps for shared management start with information sharing, then this and that. Proper management treaties not only have but can also be stronger in the fase of a canging climate.

Species can have a cosmopolitan distribution, that is, a broader distribution, or be limited to a specific region, known as endemic [@NelsonFishesOfTheWorld]. Not surprisingly, highly migratory pelagic-oceanic species are the only group that significantly differs in terms of sharing countries (one-way ANOVA *F*(5,109), *f* = 14.67, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05) with a median of 40 nations sharing these resources (Fig. 3).

@@

Due to their large migration patterns and presence in areas beyond national jurisdiction, pelagic-oceanic species such as skipjack tuna (*Katsuwonus pelamis*), Common thresher shark (*Alopias vulpinus*), and tunas (*Thunnus sp.*) are catalogued as stradling species by FAO definitions and globally managed by Regional Fisheries Management Organizations (RFMOs) [@FAO]. Sixteen major RFMOs in the world oversee the management of 101 species, most of them fished in areas beyond national jurisdiction [@CullisSuzuki:2010fi]. We estimate that 56.4% of these species (*n* = 57) are actually shared by neighbouring RFMOs with the North Atlantic Salmon Conservation Organization (NASCO), the North East Atlantic Fisheries Commission (NASCO) and the International Commission for the Conservation of Atlantic Tunas (ICCAT) sharing over 30 species with neighbouring RFMOs (Suplemental figure RFMO_MAP).

The catch from these species comes mainly from five South Pacific Islands nations, from which three (Kiribalti, Solomon and Papua New Guinea) are members of the Pacific Nauru Agreement (PNA) (Fig. 3B). The PNA manages 30% of global skipjack tuna [@Aqorau:2018bh], estimated to produce $2 billion worth of fish every year [@https://www.ffa.int/nauru_agreement]. However, the effectiveness of RFMOs responsible for managing high migratory stocks, has been previously questioned [@CullisSuzuki:2010fi].

The median for species of all other ecosystem preferences is close to, or less than, 20, as many of these might be cosmopolitan but not necessarily highly migratory (e.g. *Solea solea*, *Panulirus argus*) (Fig. 3). However, less distributed species must not be ignored. The world of fisheries is highly connected trough larvae dispersal [@Ramesh:2019va] and so, even centralized populations could be supplying populations thousands of kilometers away [@Ramesh:2019va].

Moreover, many coastal communities, specially in developing regions, and highly dependent on fishing local fish (e.g. reef fish) for food security and improved livelihoods [@CisnerosMontemayor:2016gq; Wabnitz:2018gf; @https://www.sciencedirect.com/science/article/pii/S0308597X17308357]. For example, from the top five countries fishing transboundary reef - associated species, Indonesia, Malaysia and the Philippines are among the higest fish consumption per capita in the world with 20, 60 and 40 kg per year, respectiveley [@https://www.sciencedirect.com/science/article/pii/S0308597X12001315]. In the Phillipines...

@@

@@

The ENM-Nereus consists in a multimodel approach based on a Bioclim and a Bososted Regression Tree model [@Thuiller:2009gp], a Maxent model [@Phillips:2006ff], and a Non-Parametric Probabilistic Ecological Niche Model [@Beaugrand:2011fd]. Environmental variables utilized in the models included sea surface temperature, surface pH, surface oxygen concentration, and vertically integrated (0–100 m) net primary production (NPP) [@Asch:2018ca]. The ENM-Nereus employed all of the observational data-sources previously mentioned. Global environmental conditions were obtained from 1970 to 2000 and average from three Earth System Models developed by the Geophysical Fluid Dynamics Laboratory (GFDL- https://www.gfdl.noaa.gov/earth-system-model/), the Institute Pierre Simon Laplace (IPSL- www.icmc.ipsl.fr/), and the Max Planck Institute for Meteorology (MPI- www.mpimet.mpg.de/en/science/models/). See [@Asch:2018ca; @Reygondeau:2019uh] for model details.

@@

The previous models combine observational data with a series of biotic and abiotic information to determine the probability that a species will be found in a given space at a given time. However, this does not mean that the species in question will actually be there. While the models do use approaches to double-check species occurrences (e.g. ENM-Nereus uses four different species distribution algorithms and SPD-SAU undertakes validation by means of other models), we used a fourth data source to corroborate the models’ outputs. The *Sea Around Us* estimates total reconstructed catches - i.e., catches  based on all publicly available information sources and including discards, as well as unreported and illegal catches that are not included in the FAO data available for each country. Catches are also spatially allocated [@Zeller2016]. Thus, we used the Sea Around Us catch reconstruction database from 2005 to 2014 (last ten yeares of data) as the fourth dataset to estimate transboundary species and to estimate their catch contribution within EEZs.

@@

We developed a four-criteria method for determining whether or not a species was transboundary. Only species that meet all criteria were considered as transboundary species and the analysis was done only within the EEZs of coastal states. In some cases, criteria also work as indexes to measure the uncertainty in the analysis as they provided continuos index (e.g. from 0 to 1) rather than a discrete, yes or no, index . All of the analysis was done in the statistical software *R version 3.5.2 (2018-12-20)* with the pacages *tidiverse*[@], *data.table*[@], *wesanderson*[@], *janitor*[@], *spdep*[@], *sf*[@], *sp*[@], *rgdal*[@], *tools*[@], *parallel*[@], *rfishbase*[@], *zoo*[@]. All code is available at https://github.com/jepa/FishForVisa.

@@

We assume that a species was only present in a given grid-cell if it was reported in the SAU catch database. Therefore, all species that were not reported as caught in any single year between the reference years (2005 to 2014) in a given grid-cell were dropped, regardless of the *Species Index*. The assumption relies in that if a commercial species is projected within any fishing country, such species would have been fished, and thus reported at some period of time, thus validating the models.

@@

We estimated the economic contribution of transboundary species for each country using global ex-vessel price [@Tai2017]. The data includes ex-vessel price from multiple sources and a structured interpolation method (e.g. similar countries, species) to fill in data gaps [@Sumaila:2015uc]. The contribution of transboundary species was done by aggregating the catch of each species that happened within the EEZ of the countries sharing that species. For this study, we did not include catch from areas beyond national jurisdiction. Finally, we use the catch data to determine the exploitation category of each species within a EEZ. Although this method has previously used to estimate stock status [@Grainger:1996tk; @Kleisner:2011wn], the categories presented here are intended to be seen as catch trends, and not status of each species. We only assessed species within each EEZ that had at least ten years of data between the first and last reported landings with at least five consecutive data years. Three categories were drawn from the method depending on the year's catch, the maximum catch over time and the minimum catch after the catch peak (**Table XX**). Finally, we reported the predominant category over the last ten years of catch data (2005-2014).

@@

# Suplements

<!-- # Recicled text (Not to be included) -->
